---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# clbart

<!-- badges: start -->
<!-- badges: end -->

This package implements the CL-BART model, a method that models heterogeneous 
effects for a single predictor in a conditional logistic regression. Specifically,
it is designed to be used with case-crossover designs, where each observed case
is matched to itself a number of times.

## Installation

You can install the development version of clbart from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jacobenglert/clbart")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(clbart)
library(dplyr)

library(survival)
clogit(y ~ x + as.matrix(w * z) + strata(s))

summary(cco)

w <- select(cco, starts_with('W'))
x <- select(cco, starts_with('X'))
y <- cco$Y
z <- cco$Z
s <- cco$Strata

fit <- clbart(w = w, x = x, y = y, z = z, stratum = s, 
              num_trees = 20, seed = 1, iter = 5000, warmup = 2500, thin = 5,
              alpha_sigma = 1, beta_sigma = 1)

```


<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.


```{r plots, eval = FALSE}
library(tidyverse)

# Visualize Trace Plots for Fixed Effects
fit$beta |>
  data.frame() |>
  mutate(Iteration = row_number()) |>
  pivot_longer(cols = -Iteration, names_to = 'Parameter', values_to = 'Value') |>
  ggplot(aes(x = Iteration, y = Value)) +
  geom_line() +
  geom_hline(yintercept = 0, lty = 2) +
  facet_wrap(~Parameter) +
  theme_bw()

# Summarize Fixed Effects with Credible Intervals
fit$beta |>
  data.frame() |>
  mutate(Iteration = row_number()) |>
  pivot_longer(cols = -Iteration, names_to = 'Parameter', values_to = 'Value') |>
  summarise(post_mean = mean(Value),
            post_l95 = quantile(Value, 0.025),
            post_u95 = quantile(Value, 0.925),
            .by = Parameter)

# Visualize Model Diagnostics
fit[c('tree_acc_rate','beta_acc_rate','time','logLik')] |>
  bind_cols() |>
  mutate(Iteration = row_number()) |>
  pivot_longer(cols = -Iteration, names_to = 'Parameter', values_to = 'Value') |>
  ggplot(aes(x = Iteration, y = Value)) +
  geom_line() +
  facet_wrap(~Parameter, scales = 'free') +
  theme_bw()

# Visualize Tree Complexity
fit[c('avg_tree_depth','avg_num_nodes','avg_num_leaves')] |>
  bind_cols() |>
  mutate(Iteration = row_number()) |>
  pivot_longer(cols = -Iteration, names_to = 'Parameter', values_to = 'Value') |>
  ggplot(aes(x = Iteration, y = Value, color = Parameter)) +
  geom_line() +
  theme_bw()

# Visualize Tree Splitting Variables
fit$split_props |>
  data.frame() |> 
  mutate(Iteration = row_number()) |>
  pivot_longer(cols = -Iteration, names_to = 'Covariate', values_to = 'Value') |>
  summarise(post_mean = mean(Value),
            post_l95 = quantile(Value, 0.025),
            post_u95 = quantile(Value, 0.975),
            .by = Covariate) |>
  ggplot(aes(x = Covariate, y = post_mean)) +
  geom_col(fill = 'white', color = 'black') +
  geom_errorbar(aes(ymin = post_l95, ymax = post_u95), width = 0.5) +
  theme_bw()
# 
# fit$split_props |>
#   data.frame() |> 
#   mutate(Iteration = row_number()) |>
#   pivot_longer(cols = -Iteration, names_to = 'Covariate', values_to = 'Value') |>
#   ggplot(aes(x = Iteration, y = Covariate, fill = Value)) +
#   geom_tile() +
#   theme_bw()

# Calculate CLORs (Conditional Log Odds Ratios)
w_unique <- w[c(T,F,F,F,F),]
clor <- lapply(fit$forests, \(f) predict_forest(forest = f, new_data = w_unique)) |>
  do.call(what = rbind) |>
  data.frame() |>
  mutate(Iteration = row_number())

# Visualize 10 CLORs
clor |> 
  pivot_longer(cols = -Iteration, names_to = 'ID', values_to = 'CLOR') |>
  group_by(ID) |>
  group_split() |>
  sample(20) |>
  bind_rows() |>
  ggplot(aes(x = Iteration, y = CLOR, color = ID)) +
  geom_line(show.legend = FALSE) +
  geom_hline(yintercept = 0, lty = 2) +
  facet_wrap(~ID) +
  theme_bw()

# Visualize ACLOR (Average Conditional Log Odds Ratio)
clor |> 
  pivot_longer(cols = -Iteration, names_to = 'ID', values_to = 'CLOR') |>
  summarise(ACLOR = mean(CLOR), .by = Iteration) |>
  ggplot(aes(x = Iteration, y = ACLOR)) +
  geom_line() +
  theme_bw()

# Calculate Marginal Partial ACLORs
library(parallel) # to speed up computation
pACLOR <- mclapply(colnames(w), \(x) comp_pd(model = fit, model_data = w_unique, vars = x, resolution = 20), mc.cores = 6)
pACLOR_clean <- pACLOR |>
  bind_rows() |>
  pivot_longer(cols = all_of(colnames(w)), names_to = 'Variable', values_to = 'Value') |>
  filter(!is.na(Value))

# Visualize Marginal pACLORs
pACLOR_clean |>
  summarise(pACLOR = mean(pd),
            lower = quantile(pd, .05),
            upper = quantile(pd, .95),
            .by = c(Variable, Value)) |>
  ggplot(aes(x = Value, y = pACLOR)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2) +
  facet_wrap(~Variable, scales = 'free_x') +
  theme_bw()


# Partial Dependence Function
comp_pd <- function(model = NULL, model_data = NULL, new_data = NULL, vars = NULL, resolution = 5){
  
  if(is.null(new_data)){
    pd_groups <- sapply(vars, \(v) seq(min(model_data[v]), max(model_data[v]), length.out = resolution), simplify = FALSE) |>
      expand.grid()
    
    #pd_groups <- seq(min(model_data[vars]), max(model_data[vars]), length.out = 5)
    new_data <- model_data[!(names(model_data) %in% vars)] |>
      dplyr::group_by(across(everything())) |>
      dplyr::summarise(n = n(), .groups = 'drop') |>
      merge(pd_groups)
  }
  
  pd <- model$forests |>
    lapply(\(f){
      cbind(new_data, preds = predict_forest(f, new_data = new_data[!(names(new_data) == 'n')])) |>
        dplyr::group_by(across(all_of(vars))) |>
        dplyr::summarise(pd = weighted.mean(preds, n), .groups = 'drop')
    }) |>
    do.call(what = rbind)
  
  pd$Iteration <- rep(1:(nrow(pd) / nrow(pd_groups)), each = nrow(pd_groups))
  
  return(pd)
  
}
```
